Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-09 paper/paper.tex
Moved less important optimisations out of "parameterisation" section

@@

$\mathcal{L}$ is a SSA, three address instruction set.  Destination registers
are implicit and a fresh register exists for each instruction to write its
output to.  A na\"{\i}ve way to encode $\mathcal{L}$ instructions is to have an
opcode and two operands, where each operand is either a register (i.e. a program argument
or the result of a previous instruction), or an immediate constant.

@@

\subsection{Remove nops}
Many instructions in $\mathcal{L}$ are nops that do not doing anything.
For example the instruction \verb|add x 0| does nothing.  Such instructions
can be removed from any program they appear in to leave a semantically
equivalent, but shorter, program.  We can therefore be sure that nops
will never appear in any minimal program.  By adding constraints saying that
each instruction is not a nop, we can help the underlying SAT solver's
search, which reduces the runtime of the overall procedure.

\subsection{Symmetry Reduction}
There are many instructions that are equivalent to each other.  For example,
\verb|add x y| is equivalent to \verb|add y x| -- any program containing
one instruction could have it replaced by the other instruction and
keep the same semantics.  We choose a single canonical instruction to
represent all instructions in a particular equivalence class, then add
constraints saying that no non-canonical instructions appear in the program.

\subsubsection{Implementing $\mathcal{L}$ With A Stack Machine}
Three-address code tends to be less compact than code written for a stack
machine.  It therefore would seem reasonable to use a stack machine to
execute $\mathcal{L}$ programs, since we would expect to see smaller programs
(in terms of bits used in the encoding), leading to smaller SAT instances
and faster runtimes overall.

Unfortunately we saw quite the opposite effect -- runtimes when implementing the
$\mathcal{L}$-interpreter as a stack machine were much higher than the three-address
code implementation.  We believe the reason for this
effect is that the internal state of a stack based interpreter is much more complex
to analyse than an interpreter for three-address code.  This is because a
three-address code interpreter can be implemented using a single array whose cells are
each written to exactly once (each cell corresponds to a register in the program).
By contrast a stack machine interpreter has a stack as its central data structure.
It is natural to implement this as an array, but the array's contents are accessed
in an unpredictable, non-uniform manner as the stack pointer increases and decreases
over the lifetime of the program.  This leads to a \emph{larger} SAT
instance and drastically slows down the operation of the solver.

@@

TODO
Explain parameterisations of L
Explain how we search through the parameters

>>>>

\subsection{Implementation Issues}
For performance reasons, we found that adding extra constraints on the
shape of the synthesised programs sped up synthesis.  The extra constraints
fell into two categories: eliminating nops and instruction level symmetry reduction.

\subsubsection{Remove nops}
Many instructions in $\mathcal{L}$ are nops that do not doing anything.
For example the instruction \verb|add x 0| does nothing.  Such instructions
can be removed from any program they appear in to leave a semantically
equivalent, but shorter, program.  We can therefore be sure that nops
will never appear in any minimal program.  By adding constraints saying that
each instruction is not a nop, we can help the underlying SAT solver's
search, which reduces the runtime of the overall procedure.

\subsubsection{Symmetry Reduction}
There are many instructions that are equivalent to each other.  For example,
\verb|add x y| is equivalent to \verb|add y x| -- any program containing
one instruction could have it replaced by the other instruction and
keep the same semantics.  We choose a single canonical instruction to
represent all instructions in a particular equivalence class, then add
constraints saying that no non-canonical instructions appear in the program.

@@

$\mathcal{L}$ is a SSA, three address instruction set\footnote{
We experimented with implementing $\mathcal{L}$ as a stack machine, expecting
the programs to be smaller and synthesis to be faster as a result.  We saw
the opposite effect -- the more complex interpreter led to much slower synthesis.
}.  Destination registers
are implicit and a fresh register exists for each instruction to write its
output to.  A na\"{\i}ve way to encode $\mathcal{L}$ instructions is to have an
opcode and two operands, where each operand is either a register (i.e. a program argument
or the result of a previous instruction), or an immediate constant.

@@

@@
